---
title: 'US COVID-19 Prediction'
author: "David Jackson"
date: "12/26/2020 4:45PM EST"
output: 
  html_document: 
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(RSQLite)
library(ggplot2)
library(scales)
library(plotly)
theme_set(theme_linedraw() + theme(panel.grid = element_line(linetype = 'dashed')))
```

```{r,echo=FALSE}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVID-19-DB/OURWORLD.sqlite3")
# jhcum <- dbGetQuery(db,"select * from JHU")
# jhcum$date <- as.Date($date)
# jhcum <- jhcum[order(jhcum$date),]
# jhcum <- subset(jhcum,country =="US")

```

```{r, echo=FALSE}
df2 <- dbGetQuery(db,"select * from JHU where location ='United States'")
df2$date <- lubridate::ymd(df2$date)
#df1$date <- as.Date(df1$date,format="%Y-%m-%d")
df2 <- df2[order(df2$date),]
## Cum calc
# df2 <- subset(df2,location =="United States" )
df2$date <- as.Date(df2$date)
df2death <- df2 %>% filter(date>="2020-02-29") 

```

## GGPLOt2 Plots of Daily Deaths and Cases
#### Source: Our World In Data (John Hopkin's Uni CSSE)


#### Average US Covid-19 Deaths/day

```{r}
df2death$MA7 <- movingAverage(df2death$new_deaths,7,centered=TRUE)
mean((df2$new_deaths),na.rm = TRUE)
sum((df2$new_deaths),na.rm = TRUE)
```

### Washington University
#### Institute for Health Metrics and Evaluation
* "Daily deaths is the best indicator of the progression of the pandemic, although there is generally a 17-21 day lag between infection and deaths.

* Revised US COVID19 Deaths: 300,000 by Dec. 1, 2021 (Updated Aug. 6, 2020)
* Revised US COVID19 Deaths: 410,000 by Jan. 1, 2021 (Updated Sept. 5, 2020)
* **Revised US COVID19 Deaths: 363,269 by Jan. 1, 2021 (Updated Oct. 2, 2020)**
* Revised US COVID19 Deaths: 363,269 by Feb. 1, 2021 (Updated Oct. 9, 2020)
* Revised US COVID19 Deaths: 394,693 by Feb. 1, 2021 (Updated Oct. 10, 2020)
* Revised US COVID19 Deaths: 389,087 by Feb. 1, 2021 (Updated Oct. 15,2020)
* Revised US COVID19 Deaths: 385,611 by Feb. 1, 2021 (Updated Oct. 22, 2020)
* Revised US COVID19 Deaths: 399,163 by Feb. 1, 2021 (Updated Oct. 29, 2020)
* Revised US COVID19 Deaths: 399,163 by Feb. 1, 2021 (Updated Nov. 07, 2020)
* Revised US COVID19 Deaths: 438,941 by Mar. 1, 2021 (Updated Nov. 14, 2020)
* Revised US COVID19 Deaths: 470,974 by Mar. 1, 2021 (Updated Nov. 19, 2020)
* Revised US COVID19 Deaths: 538,893 by Apr. 1, 2021 (Updated Dec. 03, 2020)
* Revised US COVID19 Deaths: 502,256 by Apr. 1, 2021 (Updated Dec. 10, 2020)
* Revised US COVID19 Deaths: 561,669 by Apr.1, 2021 (Updated Dec. 17, 2020)
* Revised US COVID19 Deaths: 567,195 by Apr.1, 2021 (Updated Dec. 23, 2020)


#### US COVID-19 Deaths  Guesstimate as of Oct. 17, 2020
On Oct 10, 2020 Washington University IHME projected 363,269 US Covid19 by April 1, 2020.Would a projection  that is within  +/- 10%   still be valid?

On Dec. 21, 2020 August Comte replied,
"Quiet simply yes. However, without providing further  information its hard to say if the modeling is any good compared to other available modeling techniques. 10% (confidence interval) in variation is generally decent for predicting three months into the future. (edited) "

* Oct. 17, 2020 Fixed math error for average US Deaths per day
* Dec. 15,2020 WU IHME Model.
* My Projections = +/- 5%
* Nov 1,  2020 = 233,706 : Actual: 229,672
* Dec.1,  2020 = 261,895 : Actual: 268045 (248800 - 274989)
* Jan.1,  2021 = 291151  : Range: 276593 - 305708
* Feb.1, 2020 = 317,482 : Range: 301607 - 333356
* Mar, 1 2021 = 319,528 (Nov. 14, 2020)
* Apr. 1, 2020 = 396,154 (Dec. 3, 2020)

### US COVID19 Projected Deaths
#### Just COVID19 Deaths for Jan. 1, 2021: 363,269

```{r}
## 399163 by Jan. 1, 2021
#
(363269 - sum(df2death$new_deaths,na.rm = TRUE)) / 7  # Feb 1, 2021
sum(df2death$new_deaths,na.rm=TRUE) +  mean(df2death$new_deaths,na.rm = TRUE) *7
sum(df2death$new_deaths,na.rm=TRUE) +  mean(df2death$new_deaths,na.rm = TRUE) *37
## 
```

```{r,echo=FALSE}
ggplot(df2death) +geom_line(aes(x=date,y=MA7,col="7 Day MA"),lwd=2) + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Predicted Deaths: Jan. 1, 2021",
       subtitle="Daily Avg (blue:1096);Feb 1(Green: 4892/363269)"
       ,x="Date date",y="Daily Deaths") +
  geom_hline(yintercept=1096,lwd=1.5,col="blue") +
    geom_hline(yintercept=4892,lwd=1.5,col="green")


```



