---
title: 'US COVID-19 Prediction'
author: "David Jackson"
date: "12/05/2020"
output: 
  html_document: 
    fig_width: 10
    fig_height: 7
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message = FALSE)
library(RSQLite)
library(ggplot2)
library(scales)
library(plotly)
theme_set(theme_linedraw() + theme(panel.grid = element_line(linetype = 'dashed')))
```

```{r,echo=FALSE}
rm(list=ls())
source("../DATA/movavg.R")
db <- db <- dbConnect(RSQLite::SQLite(),dbname= "../COVIDDB/COVID.sqlite3")
jhcum <- dbGetQuery(db,"select * from JHUDATA")
jhcum$date <- as.Date(jhcum$date)
jhcum <- jhcum[order(jhcum$date),]
jhcum <- subset(jhcum,country =="US")

```

```{r, echo=FALSE}
df2 <- dbGetQuery(db,"select * from JHDEATHS")
df2$Reported <- lubridate::ymd(df2$YMD)
#df1$Reported <- as.Date(df1$Reported,format="%Y-%m-%d")
df2 <- df2[order(df2$Reported),]
## Cum calc
# df1 <- subset(df1,Countries =="USA" )
df2$Reported <- as.Date(df2$Reported)
df2death <- df2 %>% filter(Reported>="2020-02-29") 

```

```{r,echo=FALSE}
# cumm_cases <-ggplot(jhcum) +geom_line(aes(x=date,y=confirmed),lwd=3,col="red") + 
#     scale_x_date(date_breaks = "1 month") +
#   scale_y_continuous(labels = comma) +
#    theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
#   labs(title="US Cumulative Total Cases",
#        x="Date Reported",y="Cumulative Cases") 
```
```{r}
# cumm_deaths <-ggplot(jhcum) +geom_line(aes(x=date,y=deaths),lwd=3,col="blue") + 
#     scale_x_date(date_breaks = "1 month") +
#   scale_y_continuous(labels = comma) +
#    theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
#   labs(title="US Cumulative Total Deaths",x="Date Reported",y="Cumulative Deaths")
# ggplotly(cumm_cases)
# ggplotly(cumm_deaths)
```

## GGPLOt2 Plots of Daily Deaths and Cases
#### Source: John Hopkin's Uni CSSE


#### Average US Covid-19 Deaths/day

```{r}
df2death$MA7 <- movingAverage(df2death$DEATHS,7,centered=TRUE)
mean((df2$DEATHS),na.rm = TRUE)
sum((df2$DEATHS),na.rm = TRUE)
```

### Washington University
#### Institute for Health Metrics and Evaluation
* "Daily deaths is the best indicator of the progression of the pandemic, although there is generally a 17-21 day lag between infection and deaths.

* Revised US COVID19 Deaths: 300,000 by Dec. 1, 2021 (Updated Aug. 6, 2020)
* Revised US COVID19 Deaths: 410,000 by Jan. 1, 2021 (Updated Sept. 5, 2020)
* Revised US COVID19 Deaths: 363,269 by Jan. 1, 2021 (Updated Oct. 2, 2020)
* Revised US COVID19 Deaths: 363,269 by Feb. 1, 2021 (Updated Oct. 9, 2020)
* Revised US COVID19 Deaths: 394,693 by Feb. 1, 2021 (Updated Oct. 10, 2020)
* Revised US COVID19 Deaths: 389,087 by Feb. 1, 2021 (Updated Oct. 15,2020)
* Revised US COVID19 Deaths: 385,611 by Feb. 1, 2021 (Updated Oct. 22, 2020)
* Revised US COVID19 Deaths: 399,163 by Feb. 1, 2021 (Updated Oct. 29, 2020)
* Revised US COVID19 Deaths: 399,163 by Feb. 1, 2021 (Updated Nov. 07, 2020)
* Revised US COVID19 Deaths: 438,941 by Mar. 1, 2021 (Updated Nov. 14, 2020)
* Revised US COVID19 Deaths: 470,974 by Mar. 1, 2021 (Updated Nov. 19, 2020)
* Revised US COVID19 Deaths: 538,893 by Apr. 1, 2021 (Updated Dec. 03, 2020)


#### US COVID-19 Deaths  Guesstimate as of Oct. 17, 2020

* Oct. 17, 2020 Fixed math error for average US Deaths per day
* Dec. 15,2020 WU IHME Model.
* My Projections = +/- 5%
* Nov 1,  2020 = 233,706 : Actual: 229,672
* Dec.1,  2020 = 261,895 : Range: 248800 - 274989
* Jan.1,  2021 = 291151  : Range: 276593 - 305708
* Feb.1, 2020 = 317,482 : Range: 301607 - 333356
* Mar, 1 2021 = 319,528 (Nov. 14, 2020)
* Apr. 1, 2020 = 396,154 (Dec. 3, 2020)

### US COVID19 Projected Deaths
#### Just COVID19 Deaths for Jan. 1, 2021: 363,269

```{r}
## 399163 by Jan. 1, 2021
##
(363269 - sum(df2death$DEATHS)) / 26  # Feb 1, 2021
sum(df2death$DEATHS) + (mean((df2death$DEATHS)) * 26) ## Jan .1 , 2021(363269)
## 
```

```{r,echo=FALSE}
ggplot(df2death) +geom_line(aes(x=Reported,y=MA7,col="7 Day MA"),lwd=2) + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Predicted Deaths: Jan. 1, 2021",
       subtitle="Daily Avg (blue:996);Feb 1(Grenn: 3243/363269))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=996,lwd=1.5,col="blue") +
    geom_hline(yintercept=3243,lwd=1.5,col="green")

```






```{r}
## 399163 by Feb. 1, 2021
##
(399163 - sum(df2death$DEATHS)) / 59  # Feb 1, 2021
sum(df2death$DEATHS) + (mean((df2death$DEATHS)) * 59) ## Feb.1 , 2021(389087)
##
## 
```


#### Just COVID19 Deaths for Feb. 1, 2021


```{r,echo=FALSE}
ggplot(df2death) +geom_line(aes(x=Reported,y=MA7,col="7 Day MA"),lwd=2) + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Predicted Deaths: Feb. 1, 2021",
       subtitle="Daily Avg (blue:996);Feb 1(Grenn: 2077/399163))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=996,lwd=1.5,col="blue") +
    geom_hline(yintercept=2037,lwd=1.5,col="green")

```

#### US COVID-19 Deaths by Mar. 1, 2021

```{r}
##
(438941 - sum(df2death$DEATHS)) / 87  # Mar 1, 2021
sum(df2death$DEATHS) + (mean((df2death$DEATHS)) * 87) # Feb.1 , 2021(389087)
##
## 
```
```{r}

ggplot(df2death) +geom_line(aes(x=Reported,y=MA7,col="7 Day MA"),lwd=2) + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Deaths Prediction: Mar. 1, 2020",
       subtitle="Daily Avg (blue:996);Feb 1(Grenn: 1839/399163))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=996,lwd=1.5,col="blue") +
    geom_hline(yintercept=1839,lwd=1.5,col="green")

```

#### US COVID19 Projected Deaths for Apr. 1, 2021

```{r}
## 538,893 by Apr. 1, 2021
##
(538893 - sum(df2death$DEATHS)) / 118  # Apr. 1, 2021
sum(df2death$DEATHS) + (mean((df2death$DEATHS)) * 118) ## Feb.1 , 2021(389087)
##
## 
```
```{r,echo=FALSE}
ggplot(df2death) +geom_line(aes(x=Reported,y=MA7,col="7 Day MA"),lwd=2) + 
    scale_x_date(date_breaks = "1 month") +
  scale_y_continuous(labels = comma) +
   theme(axis.text.x=element_text(angle =- 45, vjust = 0.5)) +
  labs(title="US Daily Daily Death by Apr. 1, 2021",
       subtitle="Daily Avg (Blue:993);Mar. 1(Green: 2202/538893))"
       ,x="Date Reported",y="Daily Deaths") +
  geom_hline(yintercept=996,lwd=1.5,col="blue") +
    geom_hline(yintercept=2203,lwd=1.5,col="green")

```


